Actuator line modeling of vertical-axis turbines

### 2018-05-05 paper.tex
Add OpenFOAM to abstract, intro, keywords per reviewer 4

% Abstract (Do not insert blank lines, i.e. \\)
\abstract{
    To bridge the gap between high and low fidelity numerical modeling tools for
    vertical-axis (or cross-flow) turbines (VATs or CFTs), an actuator line
    model (ALM) was developed and validated for both a high and a medium
    solidity vertical-axis turbine at rotor diameter Reynolds numbers $Re_D \sim
    10^6$. The ALM is a combination of classical blade element theory and
    Navier--Stokes based flow models, and in this study both $k$--$\epsilon$
    Reynolds-averaged Navier--Stokes (RANS) and Smagorinsky large eddy
    simulation (LES) turbulence models were tested. The RANS models were able to
    be run on coarse grids while still providing good convergence behavior in
    terms of the mean power coefficient, and also approximately four orders of
    magnitude reduction in computational expense compared with 3-D
    blade-resolved RANS simulations. Submodels for dynamic stall, end effects,
    added mass, and flow curvature were implemented, resulting in reasonable
    performance predictions for the high solidity rotor, more discrepancies for
    the medium solidity rotor, and overprediction for both cases at high tip
    speed ratio. The wake results showed that the ALM was able to capture some
    of the important flow features that contribute to VAT's relatively fast wake
    recovery---a large improvement over the conventional actuator disk model.
    The mean flow field was better realized with the LES, which still
    represented a computational savings of two orders of magnitude compared with
    3-D blade-resolved RANS, though vortex breakdown and subsequent turbulence
    generation appeared to be underpredicted, which necessitates further
    investigation of optimal subgrid scale modeling.
}

% Keywords
\keyword{wind turbine; VAWT; cross-flow turbine; blade element theory; dynamic
stall; lifting line}

@@

In this study we have developed an ALM for cross-flow turbines and embedded the
model inside LES and unsteady RANS simulations, the latter of which has not yet
been reported in the literature. This model was validated against experimental
data for the high solidity ($c/R=0.28$) UNH Reference Vertical-Axis Turbine
(UNH-RVAT) and the medium solidity ($c/R=0.07$--$0.12$) US Department of
Energy/Sandia National Labs Reference Model 2 (RM2) cross-flow turbine (at 1:6
scale), both of which are shown in Figure~\ref{fig:turbines}. Validation
datasets for both turbines were taken from \cite{Bachant2016-RVAT-Re-dep} and
\cite{Bachant2016-RM2-data}, respectively. The turbines were modeled at $Re_D
\sim 10^6$, at which experimentally measured performance and near-wake
characteristics were nearly Reynolds number
independent~\cite{Bachant2016-Energies,Bachant2016-RM2-paper}, and two orders of
magnitude higher than previous ALM investigations for VATs.

>>>>

% Abstract (Do not insert blank lines, i.e. \\)
\abstract{
    To bridge the gap between high and low fidelity numerical modeling tools for
    vertical-axis (or cross-flow) turbines (VATs or CFTs), an actuator line
    model (ALM) was developed and validated for both a high and a medium
    solidity vertical-axis turbine at rotor diameter Reynolds numbers $Re_D \sim
    10^6$. The ALM is a combination of classical blade element theory and
    Navier--Stokes based flow models, and in this study both $k$--$\epsilon$
    Reynolds-averaged Navier--Stokes (RANS) and Smagorinsky large eddy
    simulation (LES) turbulence models were tested using the open-source
    OpenFOAM computational fluid dynamics framework. The RANS models were able
    to be run on coarse grids while still providing good convergence behavior in
    terms of the mean power coefficient, and also approximately four orders of
    magnitude reduction in computational expense compared with 3-D
    blade-resolved RANS simulations. Submodels for dynamic stall, end effects,
    added mass, and flow curvature were implemented, resulting in reasonable
    performance predictions for the high solidity rotor, more discrepancies for
    the medium solidity rotor, and overprediction for both cases at high tip
    speed ratio. The wake results showed that the ALM was able to capture some
    of the important flow features that contribute to VAT's relatively fast wake
    recovery---a large improvement over the conventional actuator disk model.
    The mean flow field was better realized with the LES, which still
    represented a computational savings of two orders of magnitude compared with
    3-D blade-resolved RANS, though vortex breakdown and subsequent turbulence
    generation appeared to be underpredicted, which necessitates further
    investigation of optimal subgrid scale modeling.
}

% Keywords
\keyword{wind turbine; VAWT; cross-flow turbine; blade element theory; dynamic
stall; lifting line; OpenFOAM}

@@

In this study we have developed an ALM for cross-flow turbines and embedded the
model inside LES and unsteady RANS simulations, the latter of which has not yet
been reported in the literature. This model was implemented within the
open-source OpenFOAM CFD framework and validated against experimental data for
the high solidity ($c/R=0.28$) UNH Reference Vertical-Axis Turbine (UNH-RVAT)
and the medium solidity ($c/R=0.07$--$0.12$) US Department of Energy/Sandia
National Labs Reference Model 2 (RM2) cross-flow turbine (at 1:6 scale), both of
which are shown in Figure~\ref{fig:turbines}. Validation datasets for both
turbines were taken from \cite{Bachant2016-RVAT-Re-dep} and
\cite{Bachant2016-RM2-data}, respectively. The turbines were modeled at $Re_D
\sim 10^6$, at which experimentally measured performance and near-wake
characteristics were nearly Reynolds number
independent~\cite{Bachant2016-Energies,Bachant2016-RM2-paper}, and two orders of
magnitude higher than previous ALM investigations for VATs.
